Quantitative, Multispecies Monitoring at a Continental Scale

### 2025-02-08 DAG.tex
Updates from Overleaf

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[node distance=1.5cm]
        % Nodes
\node (qPCR_st) [node, fill=myBlue] at (0,0) {
    $\begin{array}{c}
        Z_{ik} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 \cdot K_{ik} \cdot \phi) \right) \\
        Y_{ik} \sim \mathrm{Normal} (\mu_{ik}, \sigma_{ik}) \quad \textnormal{if } Z_{ik} = 1 \\
        \mu_{ik} = \beta0_{ip} + \beta1_{ip} \cdot \ln K_{ik} \\
        \sigma_{ik} = \exp(\gamma0 + \gamma1 \cdot \ln K_{ik})
    \end{array}$
};
\node [above=of qPCR_st, yshift=-1.5cm, font=\large\bfseries] {qPCR Standard model};
\node (qPCR_en) [node, right=of qPCR_st, fill=myBlue]  at (4,0) {
    $\begin{array}{c}
        Z_{ijdbtr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 \cdot C_{ijdbt} \cdot \phi) \right) \\
        Y_{ijdbtr} \sim \mathrm{Normal} (\mu_{ijdbt}, \sigma_{ijdbt}) \quad \textnormal{if } Z_{ijr} = 1 \\
        \mu_{ijdbt} = \beta0_{ip} + \beta1_{ip} \cdot \ln C_{ijdbt} \\
        \sigma_{ijdbt} = \exp(\gamma0 + \gamma1 \cdot \ln C_{ijdbt})
    \end{array}$
};
\node [above=of qPCR_en, yshift=-1.5cm, font=\large\bfseries] {qPCR Unknown model};
\node (covariates) [node, below=of qPCR_en, fill=myBlue] { 
% + \delta0_{ijdbt} \cdot
% + \delta1_{ijdb}
    $\begin{array}{c}
        \ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        \ln D_{ijdb} = \psi_i + \ln D_{ijd} + \rho_{ijdbt}
    \end{array}$
};
\node [above=of covariates, yshift=-1.5cm, xshift=-1.3cm, font=\large\bfseries] {Covariates};
\node (metabarcoding) [node, below=of covariates, fill=myBlue] {
    $\begin{array}{c}
        R_{ijdb} \sim \mathrm{Multinomial} \left( \pi_{ijdb}, T_{jdb} \right) \\[8pt]
        \pi_{ijdb} = \frac{e^{\nu_{ijdb}}}{\sum_{i=1}^{I} e^{\nu_{ijdb}}} \\[8pt]
        \nu_{ijdb} = (\ln D_{ijdb} - \ln D_{jdb, i=\mathrm{ref}}) + NPC_{R} \cdot \alpha_i
    \end{array}$
};
\node [above=of metabarcoding, yshift=-1.5cm, xshift=-2.5cm, font=\large\bfseries] {Metabarcoding model};
\node (mock) [node, right=of metabarcoding, fill=yellow!50] {
    $\begin{array}{c}
        R_{im} \sim \mathrm{Multinomial} \left( \pi_{im}, T_{m} \right) \\[8pt]
        \pi_{im} = \frac{e^{\nu_{im}}}{\sum_{i=1}^{I} e^{\nu_{im}}} \\[8pt]
        \nu_{im} = (\ln D_{im} - \ln D_{m, i=\mathrm{ref}}) + NPC_{R} \cdot \alpha_i
    \end{array}$
};
\node [above=of mock, yshift=-1.5cm, font=\large\bfseries] {Mock metabarcoding model};
\node (spatial) [node, below=of metabarcoding, fill=green!40] {
    $\begin{array}{c}
        \ln D_{xyz} = \tau + \eta(s) + \varepsilon_{d}(s)
    \end{array}$
};
\node [above=of spatial, yshift=-1.5cm, xshift=-1.6cm, font=\large\bfseries] {Spatial model};

@@

% Legend: Latent Parameters
\node (parameters) [rectangle, draw, align=left, right=of qPCR_en, text width=12cm, fill=myGrey] at (11.9,-0.8){
    \textbf{Latent parameters} \\
    $\psi$ = mean hake eDNA concentration \\
    %$\delta0, \delta1$ = offset from mean hake \\
    $\rho$ = biological replicate random effect \\
    $\omega$ = ethanol wash effect \\
    $\xi$ = volume and dilution offset \\
    $\mu$ = mean Ct values of qPCR \\
    $\phi$ = probability of positive qPCR amplification \\
    $\sigma$ = standard deviation of qPCR Ct values \\
    $\beta0, \beta1$ = intercept and slope between $\mu$ and eDNA concentration (K and C) \\
    $\gamma0, \gamma1$ = intercept and slope between $\sigma$ and eDNA concentration (K and C) \\
    $\alpha$ = amplification efficiency relative to reference species \\
    $\pi$ = sequence proportions at the end of PCR [0,1] \\
    $\nu$ = sequence proportions at the end of PCR \\
    $\tau$ = scalar intercept\\
    $\eta$ = spatial field \\
    $\varepsilon$ =  depth specific spatial field
};
% Legend: Subscripts
\node (subscripts) [rectangle, draw, align=left, below=of qPCR_st, text width=7cm, fill=myGrey] at (0.4,0) {
    \textbf{Subscripts} \\
    $i$ = species \\
    $j$ = station \\
    $d$ = depth \\
    $b$ = biological replicate \\
    $t$ = aliquot replicate \\
    $r$ = technical replicate \\
    $p$ = qPCR plate \\
    $m$ = mock community aliquote sample \\
    $k$ = qPCR standard aliquote sample
};
% Legend: Data
\node (data) [rectangle, draw, align=left, below=of subscripts, text width=7cm, fill=myGrey] at (0.4,-5.6){
    \textbf{Data} \\
    $Z$ = qPCR amplification yes/no (0,1) \\
    $Y$ = qPCR Ct values \\
    $K$ = Known concentration in copies/$\mu$L \\
    $R$ = metabarcoding sequencing reads \\
    $T$ = total number sequencing reads
};
% Legend: State variables
\node (state) [rectangle, draw, align=left, right=of qPCR_en, text width=6cm, fill=myGrey] at (11.9,-9){
    \textbf{State variables} \\
    $C$ = unknown eDNA concentration \\
    $D$ = unknown eDNA concentration
};
    \end{tikzpicture}
    %\caption{Multifish DAG}
\end{figure}

>>>>

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[node distance=1.5cm]
        % Nodes
\node (qPCR_st) [node, fill=myBlue] at (0,0) {
    $\begin{array}{c}
        Z_{ik} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 \cdot K_{ik} \cdot \phi) \right) \\
        Y_{ik} \sim \mathrm{Normal} (\mu_{ik}, \sigma_{ik}) \quad \textnormal{if } Z_{ik} = 1 \\
        \mu_{ik} = \beta0_{ip} + \beta1_{ip} \cdot \ln K_{ik} \\
        \sigma_{ik} = \exp(\gamma0 + \gamma1 \cdot \ln K_{ik})
    \end{array}$
};
\node [above=of qPCR_st, yshift=-1.5cm, font=\large\bfseries] {qPCR Standard model};
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_st.south east) {\textbf{(2)}};

\node (qPCR_en) [node, right=of qPCR_st, fill=myBlue]  at (4,0) {
    $\begin{array}{c}
        Z_{ijdbtr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 \cdot C_{ijdbt} \cdot \phi) \right) \\
        Y_{ijdbtr} \sim \mathrm{Normal} (\mu_{ijdbt}, \sigma_{ijdbt}) \quad \textnormal{if } Z_{ijr} = 1 \\
        \mu_{ijdbt} = \beta0_{ip} + \beta1_{ip} \cdot \ln C_{ijdbt} \\
        \sigma_{ijdbt} = \exp(\gamma0 + \gamma1 \cdot \ln C_{ijdbt})
    \end{array}$
};
\node [above=of qPCR_en, yshift=-1.5cm, font=\large\bfseries] {qPCR Unknown model};
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_en.south east) {\textbf{(3)}};

\node (covariates) [node, below=of qPCR_en, fill=myBlue] { 
% + \delta0_{ijdbt} \cdot
% + \delta1_{ijdb}
    $\begin{array}{c}
        \ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        \ln D_{ijdb} = \psi_i + \ln D_{ijd} + \rho_{ijdbt}
    \end{array}$
};
\node [above=of covariates, yshift=-1.5cm, xshift=-1.3cm, font=\large\bfseries] {Covariates};
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt] covariates.south east) {\textbf{(4)}};

\node (metabarcoding) [node, below=of covariates, fill=myBlue] {
    $\begin{array}{c}
        R_{ijdb} \sim \mathrm{Multinomial} \left( \pi_{ijdb}, T_{jdb} \right) \\[8pt]
        \pi_{ijdb} = \frac{e^{\nu_{ijdb}}}{\sum_{i=1}^{I} e^{\nu_{ijdb}}} \\[8pt]
        \nu_{ijdb} = (\ln D_{ijdb} - \ln D_{jdb, i=\mathrm{ref}}) + NPC_{R} \cdot \alpha_i
    \end{array}$
};
\node [above=of metabarcoding, yshift=-1.5cm, xshift=-2.5cm, font=\large\bfseries] {Metabarcoding model};
\node[anchor=south east] at ([xshift=4pt, yshift=-15pt] metabarcoding.south east) {\textbf{(5)}};

\node (mock) [node, right=of metabarcoding, fill=yellow!50] {
    $\begin{array}{c}
        R_{im} \sim \mathrm{Multinomial} \left( \pi_{im}, T_{m} \right) \\[8pt]
        \pi_{im} = \frac{e^{\nu_{im}}}{\sum_{i=1}^{I} e^{\nu_{im}}} \\[8pt]
        \nu_{im} = (\ln D_{im} - \ln D_{m, i=\mathrm{ref}}) + NPC_{R} \cdot \alpha_i
    \end{array}$
};
\node [above=of mock, yshift=-1.5cm, font=\large\bfseries] {Mock-community metabarcoding model};
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] mock.south east) {\textbf{(1)}};

\node (spatial) [node, below=of metabarcoding, fill=green!40] {
    $\begin{array}{c}
        \ln D_{xyz} = \tau + \eta(s) + \varepsilon_{d}(s)
    \end{array}$
};
\node [above=of spatial, yshift=-1.5cm, xshift=-1.6cm, font=\large\bfseries] {Spatial model};
\node[anchor=south east] at ([xshift=19pt, yshift=-4pt] spatial.south east) {\textbf{(6)}};

@@

% Legend: Latent Parameters
\node (parameters) [rectangle, draw, align=left, right=of qPCR_en, text width=12cm, fill=myGrey] at (11.9,-0.8){
    \textbf{Latent parameters} \\
    $\psi$ = mean hake eDNA concentration \\
    %$\delta0, \delta1$ = offset from mean hake \\
    $\rho$ = biological replicate random effect \\
    $\omega$ = ethanol wash effect \\
    $\xi$ = volume filtered and aliquote dilution offset \\
    $\mu$ = mean Ct values of qPCR \\
    $\phi$ = probability of positive qPCR amplification \\
    $\sigma$ = standard deviation of qPCR Ct values \\
    $\beta0, \beta1$ = intercept and slope between $\mu$ and eDNA concentration (K and C) \\
    $\gamma0, \gamma1$ = intercept and slope between $\sigma$ and eDNA concentration (K and C) \\
    $\alpha$ = amplification efficiency relative to reference species \\
    $\pi$ = sequence proportions at the end of PCR (proportion [0,1]) \\
    $\nu$ = sequence proportions at the end of PCR (additive log-ratio)\\
    $\tau$ = scalar intercept\\
    $\eta$ = spatial field \\
    $\varepsilon$ =  depth specific spatial field
};

% Legend: Subscripts
\node (subscripts) [rectangle, draw, align=left, below=of qPCR_st, text width=7cm, fill=myGrey] at (0.4,-0.7) {
    \textbf{Subscripts} \\
    $i$ = species \\
    $j$ = station \\
    $d$ = depth \\
    $b$ = biological replicate \\
    $t$ = aliquot replicate \\
    $r$ = technical replicate \\
    $p$ = qPCR plate \\
    $m$ = mock community aliquote sample \\
    $k$ = qPCR standard aliquote sample
};
% Legend: Data
\node (data) [rectangle, draw, align=left, below=of subscripts, text width=7cm, fill=myGrey] at (0.4,-5.6){
    \textbf{Data} \\
    $Z$ = qPCR amplification (yes=1; no=0) \\
    $Y$ = qPCR Ct values \\
    $K$ = Known concentration in copies/$\mu$L \\
    $R$ = metabarcoding sequencing reads \\
    $T$ = total number sequencing reads
};
% Legend: State variables
\node (state) [rectangle, draw, align=left, right=of qPCR_en, text width=7.5cm, fill=myGrey] at (11.9,-9){
    \textbf{State variables} \\
    $C$ = unknown eDNA concentration (copies/$\mu$L) \\
    $D$ = unknown eDNA concentration (copies/L)
};
    \end{tikzpicture}
    %\caption{Multifish DAG}
\end{figure}

### 2025-02-08 Main_text.tex
Updates from Overleaf

\author[a,1]{Gledis Guri}
\author[b]{Owen Liu}
\author[a]{Ryan P. Kelly}
\author[a]{Megan Shaffer}
\author[b]{Kim Parsons}
\author[b]{Krista M. Nichols}
\author[a]{Pedro F. P. Brandão-Dias}
\author[b]{Andrew Olaf Shelton}

\affil[a]{School of Marine and Environmental Affairs, University of Washington, Seattle, Washington, USA}
\affil[b]{Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA}

@@

% Please add a significance statement to explain the relevance of your work
\significancestatement{Detailed information about living resources is often legally required to support management decisions, but the cost and difficulty of acquiring this information increases dramatically at large spatial scales. We present a scaleable, quantitative method for mapping genetic trace material from species of management interest, and use it to illustrate three-dimensional patterns of environmental DNA (eDNA) from fish species along the West Coast of the United States. The same technique is applicable to environmental samples of any kind and any number of species, and is therefore likely to broadly useful across natural-resources management and the life sciences more generally.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{
AOS, KMN, RPK, MS, and KP designed the research;
MS and PB did the laboratory and the bioinformatic work;
GG, OL, RPK, and AOS analysed the data;
GG, OL, RPK, and AOS analysed interpreted the results;
GG, OL, RPK, and AOS analysed wrote the manuscript
\hl{[Please check that this is correct and add/edit as needed XXX]}
}
\authordeclaration{The authors declare no competing interest.}

@@

\begin{abstract}
Molecular data from environmental samples can reflect the abundance of species’ DNA, an index immediately relevant to natural resources management at a broad geographic scale. These data commonly derive from assays designed and targeted for specific species (e.g., using qPCR or dPCR), or metabarcoding assays from more general PCR primers that amplify many species simultaneously. Multispecies analyses make efficient use of field samples and laboratory time, and speak to inherently multispecies questions of management and ecological research. However, unlike single-species techniques, metabarcoding by itself can reflect only the proportions of target-species DNA present, not their absolute quantities. Here, we combine qPCR and metabarcoding data derived from the same samples to map quantities of many fish species along the US West Coast in three dimensions, demonstrating a technique of practical importance for both management and ecology. We derive spatially explicit estimates of eDNA abundance for 12 common species of commercial and management importance and point the way to quantitative surveys of wild species using molecules alone. We find that species distribution maps derived from eDNA effectively capture species niches and spatial distributions. Notably, our analyses identified biodiversity hotspots that align with previously documented regions of ecological significance, such as the Columbia River plume and the Heceta Bank.
\end{abstract}

@@

The majority of eDNA shed into the environment dissipates from surface waters within hours \hl[(Brasseale et al., in review); add this reference; RYAN or OLE XXX]. Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most. Given that our sampling stations are spaced more than 30 km apart, the spatial patterns we report are likely to reflect genuine differences in fish abundance rather than artifacts of eDNA transport dynamics \cite{guri2023}. Vertical smoothing of eDNA signals appears minimal, with patterns at the scale of meters: shallow-water species (sardine, anchovy, herring) occur largely at the surface and 50m depth bins; \textit{Sebastes entomelas} occurs almost exclusively at the 50m depth bin; species found throughout the water column are either diel vertical migrators or species occupying different depths at various life-history stages (Fig. \ref{fig:fig3}\textit{B}); and bottom-dwelling fishes are detected primarily at the deepest point sampled on any station. Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig. S7}) and spatially smoothed maps (Fig. \ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig. S1}) reveals minimal differences, reinforcing the notion that the integration of eDNA signals across continental scale space is biologically sound. This further supports the use of eDNA to infer broad-scale spatial patterns in fish abundance while maintaining biological realism. \hl{[Check the framing of the last two sentences XXX]}

@@

\subsection*{Integrating DNA with traditional surveys}
Effective management of natural resources at a continental scale requires that we maximize the information content of existing sampling regimes. Here, we follow \cite{shelton2022} and \hl{[Liu et al. (submitted); add this reference; RYAN or OLE XXX]}, leveraging the same water samples to provide information on an additional 12 species of commercial importance. Elsewhere, the same set of samples has yielded information on the distribution of marine mammals \hl{(Valdivia et al, in revision); add this reference; RYAN or OLE XXX]}. The repeated use of a single set of samples highlights an important strength of eDNA samples; they allow researchers to revisit and re-sequence the genetic material in archived samples as new research questions arise or attention shifts to different taxa. The use of these samples, from a targeted fisheries survey for hake, highlights the efficiency for which eDNA can effectively provide survey data for tens or hundreds of additional species for which traditional surveys do not exist.

@@

\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
\centering
\includegraphics[width=18cm]{plots/ppt_figures/DAG.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
\caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species. Model compartment indicated in the same color are run jointly. Latent parameters, state variables, data and subscripts are indicated in gray. See \href{SI_Appendix.pdf}{SI Appendix} for full model breakdown.}
\label{fig:DAG}
\end{figure*}

@@

We follow \cite{shelton2022} and construct a compositional model for the metabarcoding data from 371 of the water samples analyzed for Pacific hake qPCR. We use the metabarcoding observations of the 8 mock communities to calibrate the relative amplification efficiency of the 12 focal species (plus hake, a 13$^{th}$ species). In the absence of qPCR data, this model would produce estimates of proportional DNA contributions for each water sample. But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI_Appendix.pdf}{SI Appendix Fig. S4}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}. Our model generates estimates of DNA concentration (copies/L) for each species. Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species; we refer to hake as our “reference species”. Hake was present in 86.7\% of samples via qPCR, and 67.7\% of samples via metabarcoding, making it an appropriate and important reference species. To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration. However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occur, like Pacific sardine in our case (\href{SI_Appendix.pdf}{SI Appendix Fig. S5}). \hl{[Check
the framing of the last two sentences XXX]}Samples lacking hake detections were sporadically distributed across the study area, and their omission appears not to have induced substantial bias in our estimates.

>>>>

\author[a,1]{Gledis Guri}
\author[b]{Owen Liu}
\author[a]{Ryan P. Kelly}
\author[a]{Megan R. Shaffer}
\author[b]{Kim Parsons}
\author[a,b,c]{Ana Ram\'on-Laca}
\author[b]{Krista M. Nichols}
\author[a]{Pedro F. P. Brandão-Dias}
\author[b]{Andrew Olaf Shelton}

\affil[a]{School of Marine and Environmental Affairs, University of Washington, Seattle, Washington, USA}
\affil[b]{Northwest Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA}
\affil[c]{Museo Nacional de Ciencias Naturales, Madrid, Spain}

@@

% Please add a significance statement to explain the relevance of your work
\significancestatement{Detailed information about living resources is often legally required to support management decisions, but the cost and difficulty of acquiring this information increases dramatically at large spatial scales. We present a scaleable, quantitative method for mapping genetic trace material from species of management interest, and use it to illustrate three-dimensional patterns of environmental DNA (eDNA) from fish species along the West Coast of the United States. The same technique is applicable to environmental samples of any kind and any number of species, and is therefore likely to be broadly useful in natural-resources management and in the life sciences more generally.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{
AOS, KMN, RPK, MS, and KP designed the research;
MRS and AR-L did the laboratory work; 
PB, RPK, and GG did the bioinformatics;
GG, OL, RPK, and AOS analyzed the data;
GG, OL, RPK, and AOS interpreted the results;
GG, OL, RPK, and AOS wrote the manuscript
\hl{[Please check that this is correct and add/edit as needed XXX]}
}
\authordeclaration{The authors declare no competing interest.}

@@

\begin{abstract}
Molecular data from environmental samples can reflect the abundance of species’ DNA, an index immediately relevant to natural-resources management at a broad geographic scale. These data commonly derive from assays designed and targeted for specific species (e.g., using qPCR or dPCR), or metabarcoding assays from more general PCR primers that amplify many species simultaneously. Multispecies analyses make efficient use of field samples and laboratory time, and speak to inherently multispecies questions of management and ecological research. However, unlike single-species techniques, metabarcoding by itself can reflect only the proportions of target-species DNA present, not their absolute quantities. Here, we combine qPCR and metabarcoding data derived from the same samples to map quantities of many fish species along the US West Coast in three dimensions, demonstrating a technique of practical importance for both management and ecology. We derive spatially explicit estimates of eDNA abundance for 12 common species of commercial and management importance and point the way to quantitative surveys of wild species using molecules alone. We find that species distribution maps derived from eDNA effectively capture species niches and spatial distributions. Notably, our analyses identified biodiversity hotspots that align with previously documented regions of ecological significance, such as the Columbia River plume and the Heceta Bank.
\end{abstract}

@@

The majority of eDNA shed into the environment dissipates from surface waters within hours \hl{(Brasseale et al., in review)[add this reference; RYAN or OLE XXX]}. Thus, the horizontal scale of environmental smoothing in our dataset likely spans several kilometers at most. Given that our sampling stations are spaced more than 30 km apart, the spatial patterns we report are likely to reflect genuine differences in fish abundance rather than artifacts of eDNA transport dynamics \cite{guri2023}. Vertical smoothing of eDNA signals appears minimal, with patterns at the scale of meters: shallow-water species (sardine, anchovy, herring) occur largely at the surface and 50m depth bins; \textit{Sebastes entomelas} occurs almost exclusively at the 50m depth bin; species found throughout the water column are either diel vertical migrators or species occupying different depths at various life-history stages (Fig. \ref{fig:fig3}\textit{B}); and bottom-dwelling fishes are detected primarily at the deepest point sampled on any station. Notably, the comparison between sample-specific eDNA estimation (\href{SI_Appendix.pdf}{SI Appendix, Fig. S7}) and spatially smoothed maps (Fig. \ref{fig:fig1} and \href{SI_Appendix.pdf}{SI Appendix, Fig. S1}) reveals minimal differences, reinforcing the notion that the integration of eDNA signals across continental scale space is biologically sound. This further supports the use of eDNA to infer broad-scale spatial patterns in fish abundance while maintaining biological realism. \hl{[Check the framing of the last two sentences XXX]}

@@

\subsection*{Integrating DNA with traditional surveys}
Effective management of natural resources at a continental scale requires that we maximize the information content of existing sampling regimes. Here, we follow \cite{shelton2022} and \hl{Liu et al. (submitted)[add this reference; RYAN or OLE XXX]}, leveraging the same water samples to provide information on an additional 12 species of commercial importance. Elsewhere, the same set of samples has yielded information on the distribution of marine mammals \hl{(Valdivia et al, in revision)[add this reference; RYAN or OLE XXX]}. The repeated use of a single set of samples highlights an important strength of eDNA samples; they allow researchers to revisit and re-sequence the genetic material in archived samples as new research questions arise or attention shifts to different taxa. The use of these samples, from a targeted fisheries survey for hake, highlights the efficiency for which eDNA can effectively provide survey data for tens or hundreds of additional species for which traditional surveys do not exist.

@@

\begin{figure*}[tbhp]  % Use figure* for a column-wide figure
\centering
\includegraphics[width=18cm]{plots/ppt_figures/DAG.pdf}  % Adjust width to 11.4 cm or 17.8 cm as per the template
\caption{Directed acyclic graph (DAG) of three models (orange, blue, and green) for estimating eDNA concentration (copies/L) for multiple (12) fish species. Model compartment indicated in the same color (2-5) are run jointly. Latent parameters, state variables, data and subscripts are indicated in gray. See \href{SI_Appendix.pdf}{SI Appendix} for full model breakdown.}
\label{fig:DAG}
\end{figure*}

@@

We follow \cite{shelton2022} and construct a compositional model for the metabarcoding data from 371 of the water samples analyzed for Pacific hake qPCR. We use the metabarcoding observations of the 8 mock communities to calibrate the relative amplification efficiency of the 12 focal species (plus hake, a 13$^{th}$ species). In the absence of qPCR data, this model would produce estimates of proportional DNA contributions for each water sample. But for each sample in which we have non-zero estimates of hake DNA concentration from qPCR, and non-zero estimate of hake proportional contribution from metabarcoding (\href{SI_Appendix.pdf}{SI Appendix Fig. S4}), we can expand from the DNA concentration of hake to provide an estimate of the DNA concentration of the 12 remaining focal species \cite{allan2023}. Our model generates estimates of DNA concentration (copies/L) for each species. Thus hake is a lynchpin that allows us to provide estimates of DNA concentration for all other species; we refer to hake as our “reference species”. Hake was present in 86.7\% of samples via qPCR, and 67.7\% of samples via metabarcoding, making it an appropriate and important reference species. To enhance sample coverage -- by increasing the number of non-zero estimates of link species -- and improve model accuracy, multiple single-species qPCR assays can be integrated as additional sources of DNA concentration. However, this approach must be applied thoughtfully, ensuring that selected qPCR targeted species have minimal co-occurance, like Pacific sardine in our case (\href{SI_Appendix.pdf}{SI Appendix Fig. S5}). Nevertheless, samples lacking hake detections here were sporadically distributed across the study area (\href{SI_Appendix.pdf}{SI Appendix Fig. S4}), and their omission appears not to have induced substantial bias in our estimates.
